---
title: "p8105_hw6_yl5214"
output: github_document
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(broom)
library(modelr)
library(purrr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


##problem 2
#load in the dataset
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

#bootstrapping
```{r}
# function of boot_sample
boot_sample = function(df) {
  
  sample_frac(df, replace = TRUE)
  
}
set.seed(1)
#draw 5000 sample and fit the model to each strap number
boot_results=
  tibble(strap_number=1:5000) |> 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(weather_df)),
    models = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results_e=map(models, broom::tidy),
    results_r=map(models, broom::glance)
  ) |> 
  unnest(results_e, results_r) |> 
  select(strap_number,term, estimate, r.squared) |> 
  group_by(term) |> 
   pivot_wider(
    names_from= term,
    values_from= estimate
  ) |> 
    mutate(
      log_beta=log(tmin*prcp)
    )

  # distribution of r2 
boot_results |> 
  ggplot(aes(x= r.squared))+
  geom_density()+
  labs(title='Distribution of r^2 estimates',
       x='r^2 estimate')

 # distribution of log(beta1*beta2)
boot_results |> 
  ggplot(aes(x=log_beta))+
  geom_density()+
  labs(title='Disribution of log(beta1*beta2) estimate',
       x='log(beta1*beta2) estimates')

# 95% confidence interval for r_squared
boot_results |> 
  summarize(
    ci_lower=quantile(r.squared, 0.025),
    ci_upper=quantile(r.squared,0.975)
  )

# 95% confidence interval for log(beta1*beta2)
boot_results |> 
  drop_na(log_beta) |> 
  summarize(
    ci_lower=quantile(log_beta, 0.025),
    ci_upper=quantile(log_beta,0.975)
  )

```
## interpretation: 
The distribution for log(beta1 * beta2) is concentrated within a specific negative range of the x-axis, and peak around -6, suggesting that the product of the regression coefficients is typically within the range of values from -12 to -4. This is a left skewed distribution. The values are on negative scale indicates that beta1 times beta2 are less than 1.
The distribution for R squared estimates is skewed towards higher values and peak at 0.92, suggesting that in most bootstrap samples, the predictors (tmin and prcp) have consistent and substantial contribution to the variability in the response variable. This suggests the model is a good fit.

# Problem 3
# load in dataset
```{r}
birthweight=
  read_csv('data/birthweight.csv') |> 
  janitor::clean_names() |> 
  drop_na() |> 
  mutate(
    babysex = recode(babysex, "1" = "male", "2" = "female"),
    frace = recode(frace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other", "9" = "Unknown"),
    mrace = recode(mrace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other"),
    malform = recode(malform, "0" = "absent", "1" = "present"),
    babysex = as.factor(babysex), 
    frace = as.factor(frace), 
    malform = as.factor(malform), 
    mrace = as.factor(mrace)
  )
```

```{r}
#build model
linear_mod= lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + ppbmi, data = birthweight)
summary(linear_mod)

birthweight_df=
  birthweight |> 
  add_predictions(linear_mod) |> 
  add_residuals(linear_mod)

# Plotting residuals against fitted values
  birthweight_df |> 
ggplot(aes(x = pred, y = resid)) + 
  geom_point() + 
  geom_hline(yintercept = 0, color = "red") +
  labs(
    x="Fitted Values",
    y="Residuals",
    title= "Model residuals against fitted values"
    )
```
# Description of model selection:
All the variables I choose is because those factors are most relevant to birth weight. For example, babysex is included in the model because of potential difference between male and female infants. Baby's head circumference and baby's length at birth are direct physical measurements at birth and likely strong predictors of weight.'delwt','gaweeks', 'ppbmi' are maternal factors which can influence fetal development and hence birth weight. Also, I choose the variables because of data-driven consideration. Each predictor in this model is statistical significance because its p-value is less than 0.05. The adjusted R-squared is 0.6994, which support this model is a good indication of birth weight.

#Compare the three models
```{r}
cv_df=
  birthweight|> 
  crossv_mc(n=100) |> 
  mutate(
    train=map(train, as_tibble),
    test=map(test, as_tibble)
  )
```

```{r}
cv_results=
  cv_df |>
  mutate(
    model1=map(train, \(df) lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + ppbmi, data = df)),
    model2= map(train, \(df) lm(bwt ~ blength+ gaweeks, data=df)),
    model3= map(train, \(df) lm(bwt ~ bhead+ blength+babysex+bhead*babysex+bhead*blength+babysex*blength+bhead*babysex*blength, data=df))
  ) |> 
  mutate(
    rmse_model1 = map2_dbl(model1, test, \(mod, df) rmse(mod, df)),
    rmse_model2 = map2_dbl(model2, test, \(mod, df) rmse(mod, df)),
    rmse_model3 = map2_dbl(model3, test, \(mod, df) rmse(mod, df)),
  ) 
```

```{r}
cv_results |>
  select(starts_with ('rmse')) |> 
  pivot_longer(
    everything(),
    names_to ='model_type',
    values_to ='rmse',
    names_prefix ='rmse_'
  ) |> 
  ggplot(aes(x =model_type, y=rmse))+
  geom_violin()
```

# Conclusion:
model 1, which is the model including 6 variables is the best model because it has the smallest rmse value, indicating the model's predictions are closer to the actual values.

